{\steve}: A hierarchical Bayesian model for Supernova Cosmology

### 2018-11-05 papers/methods/paper.tex
Updating authorlist and removing line numbers

\usepackage{lineno}
\linenumbers
\renewcommand{\arraystretch}{1.2}
\graphicspath{ {./figures/} }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\author{
	S.~R.~Hinton\altaffilmark{1},
	T.~M.~Davis\altaffilmark{1},
	A.~G.~Kim\altaffilmark{2},
	D.~Brout\altaffilmark{3},
	R.~Kessler\altaffilmark{4,5},
	J.~Lasker\altaffilmark{4,5},
	C.~Lidman\altaffilmark{6},
	A.~M\"oller\altaffilmark{7,6},
	M.~Sako\altaffilmark{3},
	D.~Scolnic\altaffilmark{5},
	M.~Smith\altaffilmark{8},
	R.~C.~Wolf\altaffilmark{9},
	Y.~Zhang\altaffilmark{10},
	M.~March\altaffilmark{3},
	S.~Allam\altaffilmark{10},
	J.~Annis\altaffilmark{10},
	S.~Avila\altaffilmark{11},
	E.~Bertin\altaffilmark{12,13},
	D.~Brooks\altaffilmark{14},
	D.~L.~Burke\altaffilmark{15,16},
	A.~Carnero~Rosell\altaffilmark{17,18},
	M.~Carrasco~Kind\altaffilmark{19,20},
	J.~Carretero\altaffilmark{21},
	C.~E.~Cunha\altaffilmark{15},
	L.~N.~da Costa\altaffilmark{18,22},
	C.~Davis\altaffilmark{15},
	J.~De~Vicente\altaffilmark{17},
	D.~L.~DePoy\altaffilmark{23},
	P.~Doel\altaffilmark{14},
	T.~F.~Eifler\altaffilmark{24,25},
	B.~Flaugher\altaffilmark{10},
	P.~Fosalba\altaffilmark{26,27},
	J.~Frieman\altaffilmark{10,5},
	J.~Garc\'ia-Bellido\altaffilmark{28},
	E.~Gaztanaga\altaffilmark{26,27},
	D.~W.~Gerdes\altaffilmark{29,30},
	R.~A.~Gruendl\altaffilmark{19,20},
	J.~Gschwend\altaffilmark{18,22},
	G.~Gutierrez\altaffilmark{10},
	W.~G.~Hartley\altaffilmark{14,31},
	D.~L.~Hollowood\altaffilmark{32},
	K.~Honscheid\altaffilmark{33,34},
	E.~Krause\altaffilmark{24},
	K.~Kuehn\altaffilmark{35},
	N.~Kuropatkin\altaffilmark{10},
	O.~Lahav\altaffilmark{14},
	M.~Lima\altaffilmark{36,18},
	M.~A.~G.~Maia\altaffilmark{18,22},
	J.~L.~Marshall\altaffilmark{23},
	F.~Menanteau\altaffilmark{19,20},
	R.~Miquel\altaffilmark{37,21},
	R.~L.~C.~Ogando\altaffilmark{18,22},
	A.~A.~Plazas\altaffilmark{25},
	E.~Sanchez\altaffilmark{17},
	V.~Scarpine\altaffilmark{10},
	R.~Schindler\altaffilmark{16},
	M.~Schubnell\altaffilmark{30},
	S.~Serrano\altaffilmark{26,27},
	I.~Sevilla-Noarbe\altaffilmark{17},
	M.~Soares-Santos\altaffilmark{38},
	F.~Sobreira\altaffilmark{39,18},
	E.~Suchyta\altaffilmark{40},
	G.~Tarle\altaffilmark{30},
	D.~Thomas\altaffilmark{11},
	and V.~Vikram\altaffilmark{41}
	\\ \vspace{0.2cm} (DES Collaboration) \\
}

@@

\affil{$^{7}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
\affil{$^{8}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
\affil{$^{9}$ Graduate School of Education, Stanford University, 160, 450 Serra Mall, Stanford, CA 94305, USA}
\affil{$^{10}$ Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}
\affil{$^{11}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}

@@

This ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population. Accurately characterising this underlying population, its evolution over redshift, and effects from host-galaxy environment, is one of the challenges of supernova cosmology. Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude of the supernova, traditionally in the Bessell $B$ band), $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space. Specifically, for any given supernova our map between absolute magnitude and apparent magnitude is given by the distance modulus:

\begin{equation}
\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \Delta M \cdot m + \text{other corrections}, \label{eq:standard}
\end{equation}

where $M_B$ is the mean absolute magnitude for all SN~Ia given $x_1=c=0$,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter. $\Delta M \cdot m$ refers to the host-galaxy mass correlation discussed in Section \ref{sec:hostgal}. The distance modulus $\mu_{\rm obs}$ is a product of our observations, however a distance modulus $\mu_C$ can be precisely calculated given cosmological parameters and a redshift. The `$\text{other corrections}$' term often includes bias corrections for traditional $\chi^2$ analyses. Bias corrections can take multiple forms, such as a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Kessler2017, Scolnic2017}.

@@

Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted distance modulus. The $\chi^2$ function minimised is

\begin{equation}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1}_{\mu} (\mu_{\rm obs} - \mu_C),
\end{equation}

where $C^{-1}_{\mu}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties for the DES supernova analysis). The predicted $\mu_C$ is defined as

\begin{eqnarray}
\mu_C &=& 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L &=& (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
E(z) &=& \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
\end{eqnarray}

where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$, and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.

@@

The only constant between survey populations is the absolute magnitude $\langle M_B \rangle$. We model the colour skewness and the redshift-dependent means and width of the colour and skew distributions individually for each survey. The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as

\begin{eqnarray}
P(M_B, x_1, c, z | \theta) &=& \mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \, \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1})\notag\\ &\quad\quad& \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1},
\end{eqnarray}

where $\theta = \lbrace \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c \rbrace$ for legibility.

@@

There are numerous results showing statistically significant correlations between host-galaxy environment and supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013}. The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}. The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014}, invokes a step function such that $\Delta M = \gamma \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $\gamma$ represents the size of the magnitude step. The scale of this step function varies from analysis to analysis, and is treated as a fit parameter. In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:

\begin{equation}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
\end{equation}

where $\delta(0)$ represents the correction at redshift zero, and $\delta(\infty)$ a parameter allowing the behaviour to change with increasing redshift. We take flat priors on $\delta(0)$ and $\delta(0)/\delta(\infty)$.

@@

This method of creating a secondary covariance matrix using partial derivatives is used throughout the traditional and BHM analyses. For each survey and band, we have two systematics --- the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty. A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details. This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey. With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes

\begin{equation}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C_\eta\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{equation}

@@

One large difference between traditional methods and BHM methods is that we treat selection effects by incorporating selection efficiency into our model, rather than relying on simulation-driven data corrections. We describe the probability that the events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are observed and pass cuts.  To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as

\begin{align}
\mathcal{L}(\theta; {\rm data}) &= P({\rm data} | \theta, S). \label{eq:like}
\end{align}

As our model so far describes components of a basic likelihood $P({\rm data}|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find

\begin{align}
\mathcal{L}(\theta; {\rm data}) &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
\end{align}

where the denominator represents an integral over all potential data $D$. This is derived in Appendix \ref{app:selection1} As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral. Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by Hamiltonian Monte-Carlo, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle). We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $d$ for simplicity, is given as

\begin{equation}
d = \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz, \label{eq:w1}
\end{equation}

where $P(m_B | z, \theta)$ can be expressed by translating the underlying $M_B$, $x_1$, and $c$ population to $m_B$ given cosmological parameters. A full derivation of this can be found in Appendix \ref{app:selection2}.

@@

The population $P(m_B | z, \theta)$ becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where

\begin{eqnarray}
m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
\sigma^{*2}_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
\end{eqnarray}

What then remains is determining the functional form of $P(S|m_B)$. For the treatment of most surveys, we find that the error function, which smoothly transitions from some constant efficiency down to zero, is sufficient. Formally, this gives

\begin{equation}
P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
\end{equation}

where $\Phi^c$ is the complementary cumulative distribution function and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function. The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}. However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a better analytic form, taking the form

\begin{equation}
P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}).
\end{equation}

@@

With the well-sampled redshift approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova. For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes

\begin{eqnarray}
d_{\rm CDF} &=& \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{eqnarray}

and is incorporated into our likelihood. This is illustrated in Figure \ref{fig:efficiency}. Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency. Further details on this choice are given in Section \ref{sec:simdes}.

@@

% color bias corrections for the {\celeven} scatter model are included in our cosmology model, the strengths of which are controlled by a fit parameter, $s_c$, however neither the {\gten} nor {\celeven} simulation fits constrained this parameter, and thus the majority of bias in the {\celeven} fits remain

>>>>

%\usepackage{lineno}
%\linenumbers
\renewcommand{\arraystretch}{1.2}
\graphicspath{ {./figures/} }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\author{
S.~R.~Hinton\altaffilmark{1},
T.~M.~Davis\altaffilmark{1},
A.~G.~Kim\altaffilmark{2},
D.~Brout\altaffilmark{3},
C.~B.~D'Andrea\altaffilmark{3},
R.~Kessler\altaffilmark{4,5},
J.~Lasker\altaffilmark{4,5},
C.~Lidman\altaffilmark{6},
E.~Macaulay\altaffilmark{7},
A.~M\"oller\altaffilmark{8,6},
M.~Sako\altaffilmark{3},
D.~Scolnic\altaffilmark{5},
M.~Smith\altaffilmark{9},
R.~C.~Wolf\altaffilmark{10},
S.~Allam\altaffilmark{11},
J.~Annis\altaffilmark{11},
S.~Avila\altaffilmark{7},
E.~Bertin\altaffilmark{12,13},
D.~Brooks\altaffilmark{14},
D.~L.~Burke\altaffilmark{15,16},
A.~Carnero~Rosell\altaffilmark{17,18},
M.~Carrasco~Kind\altaffilmark{19,20},
J.~Carretero\altaffilmark{21},
M.~Childress\altaffilmark{9},
C.~E.~Cunha\altaffilmark{15},
L.~N.~da Costa\altaffilmark{18,22},
C.~Davis\altaffilmark{15},
J.~De~Vicente\altaffilmark{17},
D.~L.~DePoy\altaffilmark{23},
P.~Doel\altaffilmark{14},
T.~F.~Eifler\altaffilmark{24,25},
B.~Flaugher\altaffilmark{11},
P.~Fosalba\altaffilmark{26,27},
J.~Frieman\altaffilmark{11,5},
J.~Garc\'ia-Bellido\altaffilmark{28},
E.~Gaztanaga\altaffilmark{26,27},
D.~W.~Gerdes\altaffilmark{29,30},
R.~A.~Gruendl\altaffilmark{19,20},
J.~Gschwend\altaffilmark{18,22},
G.~Gutierrez\altaffilmark{11},
W.~G.~Hartley\altaffilmark{14,31},
D.~L.~Hollowood\altaffilmark{32},
K.~Honscheid\altaffilmark{33,34},
E.~Krause\altaffilmark{24},
K.~Kuehn\altaffilmark{35},
N.~Kuropatkin\altaffilmark{11},
O.~Lahav\altaffilmark{14},
M.~Lima\altaffilmark{36,18},
M.~A.~G.~Maia\altaffilmark{18,22},
M.~March\altaffilmark{3},
J.~L.~Marshall\altaffilmark{23},
F.~Menanteau\altaffilmark{19,20},
R.~Miquel\altaffilmark{37,21},
E.~Morganson\altaffilmark{20},
R.~L.~C.~Ogando\altaffilmark{18,22},
A.~A.~Plazas\altaffilmark{25},
E.~Sanchez\altaffilmark{17},
V.~Scarpine\altaffilmark{11},
R.~Schindler\altaffilmark{16},
M.~Schubnell\altaffilmark{30},
S.~Serrano\altaffilmark{26,27},
I.~Sevilla-Noarbe\altaffilmark{17},
M.~Soares-Santos\altaffilmark{38},
F.~Sobreira\altaffilmark{39,18},
E.~Suchyta\altaffilmark{40},
G.~Tarle\altaffilmark{30},
D.~Thomas\altaffilmark{7},
V.~Vikram\altaffilmark{41},
and Y.~Zhang\altaffilmark{11}
\\ \vspace{0.2cm} (DES Collaboration) \\
}

@@

\affil{$^{7}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK}
\affil{$^{8}$ ARC Centre of Excellence for All-sky Astrophysics (CAASTRO)}
\affil{$^{9}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK}
\affil{$^{10}$ Graduate School of Education, Stanford University, 160, 450 Serra Mall, Stanford, CA 94305, USA}
\affil{$^{11}$ Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA}

@@

This ensemble of $\hat{m}_B$, $\hat{x}_1$ and $\hat{c}$ represents an observed population, which -- due to the presence of various selection effects -- may not represent the true, underlying supernova population. Accurately characterising this underlying population, its evolution over redshift, and effects from host-galaxy environment, is one of the challenges of supernova cosmology. Given some modelled underlying supernova population that lives in the redshift-dependent space $M_B$ (absolute magnitude of the supernova, traditionally in the Bessell $B$ band), $x_1$, and $c$, the introduction of cosmology into the model is simple -- it translates the underlying population from absolute magnitude space into the observed population in apparent magnitude space. Specifically, for any given supernova our map between absolute magnitude and apparent magnitude is given by the distance modulus:
\begin{equation}
\mu_{\rm obs} = m_B + \alpha x_1 - \beta c - M_B + \Delta M \cdot m + \text{other corrections}, \label{eq:standard}
\end{equation}
where $M_B$ is the mean absolute magnitude for all SN~Ia given $x_1=c=0$,  $\alpha$ is the stretch correction \citep{Phillips1993, Phillips1999}, and $\beta$ is the color correction \citep{Tripp1998} that respectively encapsulate the empirical relation that broader (longer-lasting) and bluer supernovae are brighter. $\Delta M \cdot m$ refers to the host-galaxy mass correlation discussed in Section \ref{sec:hostgal}. The distance modulus $\mu_{\rm obs}$ is a product of our observations, however a distance modulus $\mu_C$ can be precisely calculated given cosmological parameters and a redshift. The `$\text{other corrections}$' term often includes bias corrections for traditional $\chi^2$ analyses. Bias corrections can take multiple forms, such as a redshift-dependent function \citep{Betoule2014} or a 5D function of $c$, $x_1$, $\alpha$, $\beta$ and $z$ \citep{Kessler2017, Scolnic2017}.

@@

Traditional $\chi^2$ analyses such as that found in \citet{Riess1998, Perlmutter1999, Wood-Vasey2007, Kowalski2008, Kessler2009, Conley2011, Betoule2014}, minimise the difference in distance modulus between the observed distance modulus attained from equation \ref{eq:standard}, and the cosmologically predicted distance modulus. The $\chi^2$ function minimised is
\begin{equation}
\chi^2 = (\mu_{\rm obs} - \mu_C)^\dagger C^{-1}_{\mu} (\mu_{\rm obs} - \mu_C),
\end{equation}
where $C^{-1}_{\mu}$ is an uncertainty matrix that combines statistical and systematic uncertainties (see \citet{Brout18SYS} for a review of these uncertainties for the DES supernova analysis). The predicted $\mu_C$ is defined as
\begin{eqnarray}
\mu_C &=& 5 \log\left[ \frac{d_L}{10{\rm pc}} \right], \label{eq:mu0}\\
d_L &=& (1+z)\frac{c}{H_0} \int_0^z \frac{dz'}{E(z')},\\
E(z) &=& \sqrt{ \Omega_m (1+z')^3 + \Omega_k (1 + z')^2 + \Omega_\Lambda (1+z')^{3(1+w)}}
\end{eqnarray}
where $d_L$ is the luminosity distance for redshift $z$ given a specific cosmology, $H_0$ is the current value of Hubble's constant in $\kmsmpc$ and $\Omega_m$, $\Omega_k$, and $\Omega_\Lambda$ represent the energy density terms for matter, curvature and dark energy respectively.

@@

The only constant between survey populations is the absolute magnitude $\langle M_B \rangle$. We model the colour skewness and the redshift-dependent means and width of the colour and skew distributions individually for each survey. The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as
\begin{eqnarray}
P(M_B, x_1, c, z | \theta) &=& \mathcal{N}(M_B|\langle M_B \rangle, \sigma_{M_B}) \, \mathcal{N}(x_1 | \langle x_1(z) \rangle, \sigma_{x_1})\notag\\ &\quad\quad& \mathcal{N}^{\rm{skew}}(c| \langle c(z) \rangle, \sigma_c, \alpha_c) \label{eq:l1},
\end{eqnarray}
where $\theta = \lbrace \langle M_B \rangle, \langle x_1(z) \rangle, \langle c(z) \rangle, \sigma_{m_B}, \sigma_{x_1}, \sigma_c, \alpha_c \rbrace$ for legibility.

@@

There are numerous results showing statistically significant correlations between host-galaxy environment and supernova properties \citep{Kelly2010, Lampeitl2010, Sullivan2010, DAndrea2011, Gupta2011, Johansson2013, Rigault2013}. The latest sample of over 1300 spectroscopically confirmed type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}. The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014}, invokes a step function such that $\Delta M = \gamma \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $\gamma$ represents the size of the magnitude step. The scale of this step function varies from analysis to analysis, and is treated as a fit parameter. In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:
\begin{equation}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
\end{equation}
where $\delta(0)$ represents the correction at redshift zero, and $\delta(\infty)$ a parameter allowing the behaviour to change with increasing redshift. We take flat priors on $\delta(0)$ and $\delta(0)/\delta(\infty)$.

@@

This method of creating a secondary covariance matrix using partial derivatives is used throughout the traditional and BHM analyses. For each survey and band, we have two systematics --- the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty. A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details. This gives thirteen global systematics shared by all surveys, plus two systematics per band in each survey. With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
\begin{equation}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C_\eta\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{equation}

@@

One large difference between traditional methods and BHM methods is that we treat selection effects by incorporating selection efficiency into our model, rather than relying on simulation-driven data corrections. We describe the probability that the events we observe are drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are observed and pass cuts.  To make this extra conditional explicit, we can write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as
\begin{align}
\mathcal{L}(\theta; {\rm data}) &= P({\rm data} | \theta, S). \label{eq:like}
\end{align}
As our model so far describes components of a basic likelihood $P({\rm data}|\theta)$, and we wish to formulate a function $P(S|{\rm data},\theta)$ that describes the chance of an event being successfully observed, we rearrange the likelihood in terms of those functions and find
\begin{align}
\mathcal{L}(\theta; {\rm data}) &= \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:main}
\end{align}
where the denominator represents an integral over all potential data $D$. This is derived in Appendix \ref{app:selection1} As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral. Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by Hamiltonian Monte-Carlo, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle). We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $d$ for simplicity, is given as
\begin{equation}
d = \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz, \label{eq:w1}
\end{equation}
where $P(m_B | z, \theta)$ can be expressed by translating the underlying $M_B$, $x_1$, and $c$ population to $m_B$ given cosmological parameters. A full derivation of this can be found in Appendix \ref{app:selection2}.

@@

The population $P(m_B | z, \theta)$ becomes $\mathcal{N}(m_B|m_B^*(z), \sigma^*_{m_B})$, where
\begin{eqnarray}
m_B^*(z) &=& \langle M_B \rangle + \mu(z) - \alpha \langle x_1(z) \rangle + \beta \langle c(z) \rangle \\
\sigma^{*2}_{m_B} &=& \sigma_{M_B}^2 + (\alpha \sigma_{x_1})^2 + (\beta \sigma_c)^2.
\end{eqnarray}
What then remains is determining the functional form of $P(S|m_B)$. For the treatment of most surveys, we find that the error function, which smoothly transitions from some constant efficiency down to zero, is sufficient. Formally, this gives
\begin{equation}
P(S|m_B) = \Phi^c(m_B | \mu_{\rm CDF}, \sigma_{\rm CDF}),
\end{equation}
where $\Phi^c$ is the complementary cumulative distribution function and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function. The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}. However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a better analytic form, taking the form
\begin{equation}
P(S|m_B) = \mathcal{N}^{{\rm Skew}}(m_B | \mu_{\rm Skew}, \sigma_{\rm Skew}, \alpha_{\rm Skew}).
\end{equation}

@@

With the well-sampled redshift approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova. For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SN~Ia} becomes
\begin{eqnarray}
d_{\rm CDF} &=& \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &=& 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad& \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellow-$z$}
\end{eqnarray}
and is incorporated into our likelihood. This is illustrated in Figure \ref{fig:efficiency}. Our corrections for the DES spectroscopic data utilise the CDF functional form, with the combined low redshift surveys being modelled with the skew normal efficiency. Further details on this choice are given in Section \ref{sec:simdes}.

@@
